# First Asignment

## Loading and preprocessing the data

```{r}
df <- read.csv("activity.csv", na = "NA")
``` 


## What is mean total number of steps taken per day?


Format the "date" variable into "date" type
```{r, echo=TRUE}
library(lubridate)
df$date <- ymd(df$date)
```

### Histogram of steps by day
```{r, echo=TRUE}
library(ggplot2)
StepsDay <- tapply(df$steps, df$date, sum, na.rm = TRUE)
g <- qplot(StepsDay, geom = "histogram")
g
```

### Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
MeanStepsDay <- mean(StepsDay)
MeanStepsDay
MedianStepsDay <- median(StepsDay)
MedianStepsDay
```

-------------------------------------------------------------------------------

## What is the average daily activity pattern?

### Make a time series plot (i.e.type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=df$steps), 
                                      by=list(interval=df$interval), FUN=mean, 
                                      na.rm=TRUE)

ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
At 8:35
```{r, echo=TRUE}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

-------------------------------------------------------------------------------

## Imputing missing values

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
TotalNA <- sum(is.na(df))
TotalNA
```

### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r, echo=TRUE}
#Substitute al NA values for "0"
index <- is.na(df)
df[index] <- 0
```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
newdf <- df
index <- is.na(newdf)
newdf[index] <- 0
newdf[index] <- 0
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r, echo=TRUE}
StepsDayNa <- tapply(newdf$steps, newdf$date, sum, na.rm = TRUE)
h <- qplot(StepsDayNa, geom = "histogram")
g
```

-------------------------------------------------------------------------------

## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r, echo=TRUE}
newdf$dateType <-  ifelse(as.POSIXlt(newdf$date)$wday 
                                        %in% c(0,6), 'weekend', 'weekday')
```

### Make a panel plot containing a time series plot (i.e.type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r, echo=TRUE}
newdf <- aggregate(steps ~ interval + dateType, data=newdf, mean)
ggplot(newdf, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```








